---
title: "General Social Survey"
author: "Adrian Cruz"
date: "Sept 10, 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```

# Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.\
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Question: Is gun ownership independent of opinion on capital punishment?

$H_0$: Gun ownership and opinion on capital punishment are not related.
$H_A$: Gun ownership and opinion on capital punishment are related.

OWNGUN \~ CAPPUN both are categorical, thus use Cat_Cat.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```
People who own a gun are more likley to favor capital punishment. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

75.68 percent of people who own a gun favor capital punishment while the other 24.32 percent that own a gun oppose of it. For those who dont own a gun 51.72 percent of them favor it while a big 70.97 dont.

### Inferential Results  

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Since the P-value=0.02, we must reject the null hypothesis since it is lower than 0.05, meaning that gun ownership and capital punishment are related. In conclusion, the odds for people who own a gun favoring capital punishments is 2.271.

# Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Is there a difference between years of education between those who voted and did not vote in 2012?

$H_0$: There is no difference in EDUC between those who voted and did not vote in 2012.
$H_A$: There is a difference in EDUC between those who voted and did not vote in 2012.

EDUC\~VOTE12 NUMERICAL\~CATEGORIAL, thus use Numerical\~Cat

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

People who voted in 2012 tended to have more education than those who did not vote in 2012. <!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

People on average who voted in 2012 had an education mean of 15.7 while those who didnt vote in 2012 had a education mean of 13.4. This shows a increase in years of education if voting in 2012 rather than not voting.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

Since the p.value(1.247e-07) is lower then 0.05, we must reject the null hypothesis, meaning that there is a difference between years of education and if you voted in 2012 or not.

# Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
